---
title: "3.3 Survival Analysis Notebook"
output: html_notebook
---

#Setup 
###packages, working directory

```{r setup}
setwd("~/Desktop/Code/Thesis")
library(tidyverse)
library(lubridate)
library(survival)
library(survminer)
library(GGally)
library(ggplot2)
library(ggfortify)
library(rlist)
library(rms)
library(broom)
library(survtools)
library(Epi)
library(gridGraphics)
library(grid)
library(gridExtra)
library(gtable)
```

#Data Loading
### load data

```{r}
load("~/Desktop/Code/Thesis/Data_for_Analysis/bfinal.Rdata")
load("~/Desktop/Code/Thesis/Data_for_Analysis/eg.rda")

eg_lor_full$ID <- as.numeric(eg_lor_full$ID)
data <- left_join(bfinal[c(1:7)], eg_lor_full[c(1, 3:4, 16, 18)], by=c("RAUMID" = "ID"))
```

### converting to long event format

```{r}
# converting to long event format -----------------------------------------
start <- as.Date("2000-12-31")
data$start_S1 <- time_length(difftime(data$gcode1yr, start), "years")
data$start_S2 <- time_length(difftime(data$gcode2yr, start), "years")
data$start_S3 <- time_length(difftime(data$gcode3yr, start), "years")
data$start_S4 <- time_length(difftime(data$gcode4yr, start), "years")
data$start_S5 <- time_length(difftime(data$gcode5yr, start), "years")
data <- dplyr::select(data, c(1:7, 12:16, 8:11))
```

### creating event intervals for time between stages 1-2, 2-3, 3-4, 4-5.
```{r}
stage1 <- data[c(1, 8:9)] %>% 
  dplyr::rename(startTime = 2, endTime = 3) %>% 
  filter(!is.na(startTime))
stage1$event <- ifelse(!is.na(stage1$endTime), c("tg2"), c("cens"))

stage2 <- data[c(1, 9:10)] %>% 
  rename(startTime = 2, endTime = 3) %>% 
  filter(!is.na(startTime))
stage2$event <- ifelse(!is.na(stage2$endTime), c("tg3"), c("cens"))

stage3 <- data[c(1, 10:11)] %>% 
  rename(startTime = 2, endTime = 3) %>% 
  filter(!is.na(startTime))
stage3$event <- ifelse(!is.na(stage3$endTime), c("tg4"), c("cens"))

stage4 <- data[c(1, 11:12)] %>% 
  rename(startTime = 2, endTime = 3) %>% 
  filter(!is.na(startTime))
stage4$event <- ifelse(!is.na(stage4$endTime), c("tg5"), c("cens"))

dataL <- rbind(stage1, stage2, stage3, stage4) %>% group_by(RAUMID)

end <- as.Date("2020-01-01")
dataL$endTime <- ifelse(
  is.na(dataL$endTime), time_length(difftime(end, start), "years"), dataL$endTime
)

dataL$event <- as.factor(dataL$event)

labels=c(0:4)
dataL$event_num <- as.numeric(as.character(factor(dataL$event, labels=c(0:4))))

rm(stage1, stage2, stage3, stage4, bfinal, eg_lor_full)
```

###creating treatment dummy variable based on whether policy was active during the event interval

```{r}
data$treatment_starttime <- time_length(difftime(data$F_IN_KRAFT.x, start), "years")

dataL <- left_join(dataL, data[c(1, 17)], by=c("RAUMID")) %>% tibble::rowid_to_column(., "ID")

dataL$treat <- ifelse(
  dataL$startTime > dataL$treatment_starttime, 1, ifelse(
    dataL$endTime > dataL$treatment_starttime, 1, 0
  )
)

dataL[, 8][is.na(dataL[, 8])] <- 0
```

### creating treatment dummy variable based on whether area was within 500 meters of a protection area during event interval

```{r}
# treatment dummy variable: method correct method 
data$treatment_starttime_500 <- time_length(difftime(data$F_IN_KRAFT.y, start), "years")

dataL <- left_join(dataL, data[c(1, 18)], by=c("RAUMID"))
dataL$treat500 <- ifelse(
  dataL$startTime > dataL$treatment_starttime_500, 1, ifelse(
    dataL$endTime > dataL$treatment_starttime_500, 1, 0
  )
)
dataL[, 10][is.na(dataL[, 10])] <- 0
```
  
# basic surv curve
## fitting
Creating basic survfit object
```{r}
mfit1 <- survfit(Surv(startTime, endTime, event) ~treat, id=ID, data=dataL)
print(mfit1)
```

##printing results stuff
```{r}
summary(mfit1)
summary(mfit1)$table
write.table((summary(mfit1)$table), file = "~/Desktop/Code/Thesis/misc_plots/survtable.txt", sep = ",", quote = FALSE, row.names = T)
```

```{r}
fit <- mfit1
d <- data.frame(time = fit$time,
                  n.risk = fit$n.risk,
                  n.event = fit$n.event,
                  n.censor = fit$n.censor,
                  surv = fit$pstate,
                  upper = fit$upper,
                  lower = fit$lower
                  )
names(fit)
```

## plotting
#### Epi: makes facets 
```{r}
# [MUST COPY PASTE INTO CONSOLE]
# legend("top", c("untreated", "treated"), col=c("red", "blue"), pch=0, pt.cex = 0.5, cex = .7, box.lty=1, box.lwd=1, box.col="black")
plotCIF(mfit1, col = c("red", "blue"),
        event=1, 
        main = "Progression Stage 1 -> 2", 
        xlab = "Time (Years)",
        ylab = "Cumulative Incidence", 
        ylim = c(0, 1))
grid.echo()
a <- grid.grab()

p2 <- plotCIF(mfit1, col = c("red", "blue"),
        event=2, 
        main = "Progression Stage 2 -> 3", 
        xlab = "Time (Years)",
        ylab = "Cumulative Incidence", 
        ylim = c(0, .5))
grid.echo()
b <- grid.grab()

p3 <- plotCIF(mfit1, col = c("red", "blue"),
        event=3, 
        main = "Progression Stage 3 -> 4", 
        xlab = "Time (Years)",
        ylab = "Cumulative Incidence", 
        ylim = c(0, .2))
grid.echo()
c <- grid.grab()

p4 <- plotCIF(mfit1, col = c("red", "blue"),
        event=4, 
        main = "Progression Stage 4 -> 5", 
        xlab = "Time (Years)",
        ylab = "Cumulative Incidence", 
        ylim = c(0, .1))
grid.echo()
d <- grid.grab()
```
```{r}
z <- rbind(a, b, c, d)
z$widths <- unit.pmax(a$widths, b$widths, c$widths, d$widths)
```

#### [ugly] base
```{r}
plot(mfit1, col=c(1,2), xscale=12, mark.time=FALSE, lwd=1, 
xlab="Years post diagnosis", ylab="Survival")
legend("topright", c("treatment=0", "treatment=1"), col=1:2, lwd=2, bty='n')
```

```{r}
plot(mfit1,
     mark.time=FALSE,
     lwd=2, 
     xscale=12, 
     col = topo.colors(4), 
     xlab="Years post in stage x",
     ylab="Probability of being in state x according to time")
legend("topright", col=n.risk, legend=c("Stage 1-2", "Stage 2-3", "Stage 3-4", "Stage 4-5"))
```


####competing risks
```{r}
ggcompetingrisks(mfit1)
```

#### [don't work] fancy `survminer` plots
```{r}
tidy_null <- tidy(nullsurv)

```


```{r}
print(mfit2, rmean=240, scale=12)
survminer::ggcompetingrisks(nullsurv)

surv_summary(nullsurv)

test <- coxph(Surv(startTime, endTime, event) ~ as.factor(treatedduring), id=RAUMID, data=dataL)

survminer::ggcoxdiagnostics(test)
ggcoxfunctional(test, data=dataL, id=RAUMID)
survminer::ggcox
ggcoxzph(cox.zph(test))
```


# Cox model

## fitting
```{r}
cfit <- coxph(Surv(startTime, endTime, event) ~ treat, id=ID, data=dataL)
cfit
summary(cfit)
cfit$coefficients
```

```{r}
HR <- round(exp(coef(cfit)), 2)
CI <- round(exp(confint(cfit)), 2)
P <- round(coef(summary(cfit))[,5], 3)
# Names the columns of CI
colnames(CI) <- c("Lower", "Higher")
# Bind columns together as dataset
table2 <- as.data.frame(cbind(HR, CI, P))
table2
```

##plot

### base
```{r}
plot(cfit)
```



