---
title: "#R Script for SOCI:20003 Final"
output: html_document
code_folding: show
toc: true
toc_float: true
number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exploratory Analysis of real estate value and demographic change in Berlin 2003-2011

### Setting up environment

#### Loading necessary packages
```{r message = FALSE}
library(tidyverse)
library(httr)
library(sf)
library(dplyr)
library(tmap)
library(rgdal)
library(geojsonsf)
library(raster)
library(speciesRaster)
library(fasterize)
library(ggstatsplot)

setwd("~/Desktop/Code/Thesis")
```

#### Creating functions to download data from FIS Broker (background setup)
```{r class.source = 'fold-hide'}
get_X_Y_coordinates <- function(x) {
  sftype <- as.character(sf::st_geometry_type(x, by_geometry = FALSE))
  if(sftype == "POINT") {
    xy <- as.data.frame(sf::st_coordinates(x))
    dplyr::bind_cols(x, xy)
  } else {
    x
  }
}

sf_fisbroker <- function(url) {
  typenames <- basename(url)
  url <- httr::parse_url(url)
  url$query <- list(service = "wfs",
                    version = "2.0.0",
                    request = "GetFeature",
                    srsName = "EPSG:25833",
                    TYPENAMES = typenames)
  request <- httr::build_url(url)
  print(request)
  out <- sf::read_sf(request)
  out <- sf::st_transform(out, 3035)
  out <- get_X_Y_coordinates(out)
  out <- st_as_sf(as.data.frame(out))
  return(out)
}

export_format <- c(
  "geojson", 
  "sqlite"
)

sf_save <- function(z, fname) {
  ifelse(!dir.exists(fname), dir.create(fname), "Folder exists already")
  ff <- paste(file.path(fname, fname), export_format, sep = ".")
  purrr::walk(ff, ~{ sf::st_write(z, .x, delete_dsn = TRUE)})
  saveRDS(z, paste0(file.path(fname, fname), ".rds"))
}
```

#### Loading Data

##### Shapefiles
```{r}
#Berlin polygon
load(url("https://userpage.fu-berlin.de/soga/300/30100_data_sets/berlin_district.RData"))
berlin.sf <- st_transform(berlin.sf, 3035)

#Verkehrszellen (2003 file)
vz <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_vz") 
vz <- vz %>% separate(gml_id, c("s_vz.", "gml_id"), sep=5) %>% 
  dplyr::select("gml_id", "geometry")
vz$s_vz. <- NULL

#LOR areas (2011 file)
lor <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_lor_plan")
lor <- lor %>% 
  separate(gml_id, c("s_lor_plan", "gml_id"), sep=11) %>%
  dplyr::select("gml_id", "geometry")
```

##### Real estate data
Downloading from Berlin's open data portal using their API; subsetting and cleaning
```{r}
brw_processing <- function(x){
  y <- x %>%
    dplyr::filter(NUTZUNG != "G - Gewerbe") %>%
    dplyr::select(gml_id, BRW, geometry) %>%
    filter(!st_is_empty(.)) 
}

brw_outliers <- function(x){
  val <- x$BRW
  yUL <- val %>% quantile(.97, na.rm = TRUE)
  
  out <- x %>%
    dplyr::filter(BRW < yUL)
  return(out)
}

brw2012 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2012")
brw2011 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2011")
brw2010 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2010")

brw2004 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2004")
brw2003 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2003") 
brw2002 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2002") 

brw2012 <- brw_processing(brw2012) %>% brw_outliers()
brw2011 <- brw_processing(brw2011) %>% brw_outliers()
brw2010 <- brw_processing(brw2010) %>% brw_outliers()

brw2004 <- brw_processing(brw2004) %>% brw_outliers()
brw2003 <- brw_processing(brw2003) %>% brw_outliers()
brw2002 <- brw_processing(brw2002) %>% brw_outliers()
```

##### Demographic Data
Loading and cleaning demographic data (already partially cleaned in Excel)
```{r}
mss_2003 <- read.csv("~/Desktop/Code/Thesis/demographic/2003_MSS_cut.csv")
mss_2003 <- mss_2003 %>% separate(Verkehrszellen,
                                  c("VKZ_num", "VKZ_name"), 
                                  sep = 4)
mss_2003 <- mss_2003 %>% 
  dplyr::select("VKZ_num",
                "VKZ_name",
                "Langzeit.arbeitslose.über.1.Jahr.am.31.12.02.pro.100.EW.18.60.J.", 
                "Arbeitslose.insgesamt.31.12.02.pro.100.EW.18.60.J.",
                "Deutsche.Sozialhilfe.Empfänger.pro.100.deutsche.EW",
                "Über.64.Jährige.pro.100.EW",
                "Ausländer.pro.100.EW",
                "EW.aus.EU.Staaten.pro.100.EW") %>%
  rename(long_unemp2003 = Langzeit.arbeitslose.über.1.Jahr.am.31.12.02.pro.100.EW.18.60.J.,
         unemp2003 = Arbeitslose.insgesamt.31.12.02.pro.100.EW.18.60.J.,
         gWelfare2003 = Deutsche.Sozialhilfe.Empfänger.pro.100.deutsche.EW,
         seniors2003 = Über.64.Jährige.pro.100.EW,
         foreign2003 = Ausländer.pro.100.EW,
         european2003 = EW.aus.EU.Staaten.pro.100.EW)
mss_2003[,3:8] <- lapply(mss_2003[,3:8], as.numeric)
mss_2003[339,1] <- c("Berlin")
mss_2003[339,2] <- c("Berlin")

##### MSS 2011
mss_2011 <- read.csv("~/Desktop/Code/Thesis/demographic/2011_MSS_cut.csv")
mss_2011 <- mss_2011 %>% 
  dplyr::select("Raumid",
                "Gebiet",
                "Status3", 
                "Status1",
                "E22",
                "E6",
                "E7",
                "E18") %>%
  rename(long_unemp2011 = Status3,
         unemp2011 = Status1,
         gWelfare2011 = E22,
         seniors2011 = E6,
         foreign2011 = E7,
         european2013 = E18)
mss_2011[,3:8] <- lapply(mss_2011[,3:8], as.numeric)
```

#### Rasterizing Data

##### Real Estate Data

Prerasterization: setup functions are used to reduce code; the first function creates a raster from the given sf object and the second smooths the data (some parcels are overly valued)
```{r}
#template raster for Berlin
berlin_template <- raster(extent(berlin.sf), 
                          resolution = 216, 
                          crs = st_crs(berlin.sf)$proj4string)

#function to create a raster from file
brwrast <- function(x){
  out <- fasterize(x, berlin_template, field = "BRW", background = NA)
  return(out)
}

#function to smooth data
brwfocal <- function(x){
  fweight <- focalWeight(x, d=3, type="Gauss")
  out <- focal(x, w=fweight, fun = "mean", NAonly=TRUE)
  return(out)
}
```

Rasterizing files based on functions
```{r}
brw2012rast <- brw2012 %>% brwrast() %>% brwfocal()
brw2011rast <- brw2011 %>% brwrast() %>% brwfocal()
brw2010rast <- brw2010 %>% brwrast() %>% brwfocal()

brw2004rast <- brwrastbrw2004 %>% brwrast() %>% brwfocal()
brw2003rast <- brwrastbrw2003 %>% brwrast() %>% brwfocal()
brw2002rast <- brwrastbrw2002 %>% brwrast() %>% brwfocal()
```

Consolidating files:
The result of the rasterization is our first variable to analyze: the change in average parcel valuation from 2003 to 2011. The map displays this raster. 
```{r}
brw2011ave <- stack(brw2012rast, brw2011rast, brw2010rast) %>% 
  approxNA(method = "linear", rule = 2)
brw2011ave <- overlay(brw2011ave, fun = "mean")

brw2003ave <- stack(brw2004rast, brw2003rast, brw2002rast) %>% 
  approxNA(method = "linear", rule = 2)
brw2003ave <- overlay(brw2003ave, fun = "mean") 

brwChange <- brw2011ave - brw2003ave

tmap_mode("view")
berlinbase <- tm_shape(berlin.sf) + tm_fill(alpha = .4)
testmap <- berlinbase + tm_shape(brwChange) + tm_raster(alpha = .8)
testmap
```

##### Demographic data

First, the data is spatialized via a join to the correct shapefile:
```{r}
mss_2003sf <- left_join(mss_2003, vz, by =  c("VKZ_num" = "gml_id")) %>% 
  st_as_sf() %>% st_transform(3035) %>% na.omit() %>% filter(!st_is_empty(.))

mss_2011sf <- left_join(mss_2011, lor, by =  c("Raumid" = "gml_id")) %>%
  st_as_sf() %>% 
  st_transform(3035) %>% 
  filter(!st_is_empty(.))
```

Then, a separate raster is created for each variable for each year
```{r}
unemp2003r <- rasterize(mss_2003sf, berlin_template, field = "unemp2003")
longunemp2003r <- rasterize(mss_2003sf, berlin_template, field = "long_unemp2003")
gWelfare2003r <- rasterize(mss_2003sf, berlin_template, field = "gWelfare2003")
seniors2003r <- rasterize(mss_2003sf, berlin_template, field = "seniors2003")
foreign2003r <- rasterize(mss_2003sf, berlin_template, field = "long_unemp2003")
eu2003r <- rasterize(mss_2003sf, berlin_template, field = "long_unemp2003")

unemp2011r <- rasterize(mss_2011sf, berlin_template, field = "unemp2011")
longunemp2011r <- rasterize(mss_2011sf, berlin_template, field = "long_unemp2011")
gWelfare2011r <- rasterize(mss_2011sf, berlin_template, field = "gWelfare2011")
seniors2011r <- rasterize(mss_2011sf, berlin_template, field = "seniors2011")
foreign2011r <- rasterize(mss_2011sf, berlin_template, field = "long_unemp2011")
eu2011r <- rasterize(mss_2011sf, berlin_template, field = "long_unemp2011")
```

The above rasters are then subtracted from each other to view the change between 2003 and 2011
```{r}
unempChange <- unemp2011r - unemp2003r
longunempChange <- longunemp2011r - longunemp2003r
gWelfareChange <- gWelfare2011r - gWelfare2003r
seniorsChange <- seniors2011r - seniors2003r
foreignChange <- foreign2011r - foreign2003r
euChange <- eu2011r - eu2003r
```

```{r}

```

```{r}

```





```{r}

```
