---
title: "Thesis: Berlin Gentrification Data Preparation — Property Value Data"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up environment

### Loading necessary packages
```{r message = FALSE}
library(automap)
library(tidyverse)
library(httr)
library(raster)
library(sf)
library(tmap)
library(rgdal)
library(geojsonsf)
library(dismo)
```

### Creating functions to download data from FIS Broker (background setup)
```{r class.source = 'fold-hide'}
get_X_Y_coordinates <- function(x) {
  sftype <- as.character(sf::st_geometry_type(x, by_geometry = FALSE))
  if(sftype == "POINT") {
    xy <- as.data.frame(sf::st_coordinates(x))
    dplyr::bind_cols(x, xy)
  } else {
    x
  }
}

sf_fisbroker <- function(url) {
  typenames <- basename(url)
  url <- httr::parse_url(url)
  url$query <- list(service = "wfs",
                    version = "2.0.0",
                    request = "GetFeature",
                    srsName = "EPSG:25833",
                    TYPENAMES = typenames)
  request <- httr::build_url(url)
  print(request)
  out <- sf::read_sf(request)
  out <- sf::st_transform(out, 3035)
  out <- get_X_Y_coordinates(out)
  out <- st_as_sf(as.data.frame(out))
  return(out)
}
```

### Loading Data

Downloading from Berlin's open data portal using their API; subsetting and cleaning.

Downloading the data: 
```{r eval = FALSE}
brw2004 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2004")
brw2003 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2003") 
brw2002 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2002")
load(url("https://userpage.fu-berlin.de/soga/300/30100_data_sets/berlin_district.RData"))
berlin_base <- st_transform(berlin.sf, 3035)
```

```{r echo=FALSE, warning=FALSE, results='hide'}
brw2004 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2004.shp")
brw2003 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2003.shp")
brw2002 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2002.shp")
berlin_base <- readOGR("~/Desktop/Code/Thesis/shapefiles/berlin_sf.shp")
```

Processing the data: 
```{r warning=FALSE}
# functions to make processing Bodenrichtwert data easier
brw_processing <- function(x){
  # casts objects from SpatialPolygons to sf dataframe
  y <- x %>% st_as_sf() %>%
    st_transform(3035) %>%
    # filters out only residential areas,
    dplyr::filter(NUTZUNG == "W - Wohngebiet") %>%
     # only selects relevant columns
    dplyr::select(gml_id, BRW, geometry) %>%
     # removes empty geometries
    filter(!st_is_empty(.))
  # removes outliers from data beyond the .95 threshold on a normal distribution
  val <- y$BRW
  yUL <- val %>% quantile(.95, na.rm = TRUE)
  out1 <- y %>%
    dplyr::filter(BRW < yUL)
  # casts the geometry of the dataframe from polygons to points for easier interpolation
  out <- st_centroid(out1)
  return(out)
}

# this really should be with fewer lines using lapply();
# as such, under construction.
brw2012 <- brw_processing(brw2012)
brw2011 <- brw_processing(brw2011)
brw2010 <- brw_processing(brw2010)

brw2003comb <- rbind(brw2002, brw2003, brw2004) %>% as_Spatial()
```

## Rasterizing Data

Once loaded and processed, we begin rasterizing this data. This process relies extremely heavily upon Hiljmans (2016), found at: https://rspatial.org/raster/analysis/4-interpolation.html, both in terms of methods of rasterization and the evaluation of each method. The steps are outlined below: 

### Evaluation Functions
Creating template raster
```{r}
berlin_template <- raster(extent(berlin_base), 
                          resolution = 424, 
                          crs = st_crs(berlin_base)$proj4string)
```

### Method 1: Proximity Polygons

This is a very basic method of interpolation that I'll use to compare to more advanced methods.
```{r warning=FALSE}
berlin_v <- voronoi(brw2003comb)
vberlin <- raster::intersect(berlin_v, raster::aggregate(berlin_base))
b_pp <- rasterize(vberlin, berlin_template, "BRW")
plot(b_pp)
```
This method is just to provide an example of an extremely basic method of interpolation — we can see some really dramatic cutoffs around the city that might not do a great job of reflecting the actual price difference on different blocks. Nonetheless, still useful for comparison. 

### Method 2: Kriging

Kriging is a relatively complex method of interpolation — while the specifics of how exactly kriging works is beyond both my expertise in spatial statistics and the scope of this text, it is widely used for interpolation. We use the `automap` package to cut out a lot of the variogram calculation normally used in this context.  This code automatically interpolates the parcel data, and compares each different possible variogram model, ultimately using an exponential model. 

```{r warning=FALSE, error=FALSE, results='hide'}
#unclear why this is necessary, but kriging won't complete unless the CRSs are reassigned. 
bbase <- as(berlin_template, "SpatialPixels")
crs(brw2003comb) <- crs(bbase) <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"
brw03_krigd <- autoKrige(formula=BRW~1, brw2003comb, new_data=bbase, model = c("Sph", "Exp", "Gau", "Ste", "Nug"))
```

The `autoKrige()` function returns an object can show us quite a lot about the kriging interpolation, for anyone interested.  

```{r echo=FALSE}
plot(brw03_krigd)
```

This is then converted to a raster
```{r}
brw03_pred <- brw03_krigd$krige_output
brw03_pred <- raster(brw03_pred)
brw03_pred <- mask(x = brw03_pred, mask = berlin_base)
plot(brw03_pred)
```
This looks a lot better! 

Now, to obtain the difference in real estate values between 2003 and 2011 we do the same process, but with the 2011 data; only the final output is shown. 

```{r warnings=FALSE, results='hide'}
crs(brw2011comb) <- crs(bbase) <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"
brw11_krigd <- autoKrige(formula=BRW~1, brw2011comb, new_data=bbase, model = c("Sph", "Exp", "Gau", "Ste", "Nug"))
brw11_krigd <- brw11_krigd$krige_output
brw11_krigd <- raster(brw11_krigd)
brw11_krigd <- mask(x = brw11_krigd, mask = berlin_base)
plot(brw11_krigd)
```

Next, we want to do this with all the `Bodenrichtwert` datasets from 2001-2019; the easiest way to do this is by creating a function. This function will take the datasets, process them, and create the aggregated rasters similar to the other datasets. 

```{r}
brw_calc <- function(y1, y2, y3){
  df <- rbind(y1, y2, y3) %>% as_Spatial()
  crs(df) <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"
  x <- autoKrige(formula=BRW~1, df, new_data=bbase, model = c("Sph", "Exp", "Gau", "Ste", "Nug"))
  y <- x$krige_output
  z <- raster(y)
  out <- mask(x = z, mask = berlin_base)
  return(out)
}
```

All of the Bodenrichtwert datasets are downloaded and run through this function to produce a raster for each odd year from 2003-2017. 

Downloading the data: 
```{r eval = FALSE}
brw2020 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2020") %>% brw_processing()
brw2019 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2019") %>% brw_processing()
brw2018 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2018") %>% brw_processing()
brw2017 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2017") %>% brw_processing()
brw2016 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2016") %>% brw_processing()
brw2015 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2015") %>% brw_processing()
brw2014 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2014") %>% brw_processing()
brw2013 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2013") %>% brw_processing()
brw2012 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2012") %>% brw_processing()
brw2011 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2011") %>% brw_processing()
brw2010 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2010") %>% brw_processing()
brw2009 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2009") %>% brw_processing()
brw2008 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2008") %>% brw_processing()
brw2007 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2007") %>% brw_processing()
brw2006 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2006") %>% brw_processing()
brw2005 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2005") %>% brw_processing()
brw2004 <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2004") %>% brw_processing()
```

```{r echo=FALSE, warnings=FALSE, message=FALSE}
brw2020 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2020.shp") 
brw2019 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2019.shp") 
brw2018 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2018.shp") 
brw2017 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2017.shp") 
brw2016 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2016.shp") 
brw2015 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2015.shp") 
brw2014 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2014.shp") 
brw2013 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2013.shp") 
brw2009 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2009.shp") 
brw2008 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2008.shp")
brw2007 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2007.shp")
brw2006 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2006.shp")
brw2005 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2003.shp")
brw2004 <- readOGR(dsn = "~/Desktop/Code/Thesis/brw_shapefiles/brw2005.shp")
```

```{r warnings=FALSE, message=FALSE}
brw2005_krigd <- brw_calc(brw2004, brw2005, brw2006)
brw2007_krigd <- brw_calc(brw2006, brw2007, brw2008)
brw2009_krigd <- brw_calc(brw2008, brw2009, brw2010)
brw2011_krigd <- brw_calc(brw2010, brw2011, brw2012)
brw2013_krigd <- brw_calc(brw2012, brw2013, brw2014)
brw2015_krigd <- brw_calc(brw2014, brw2015, brw2016)
brw2017_krigd <- brw_calc(brw2016, brw2017, brw2018)
brw2019_krigd <- brw_calc(brw2018, brw2019, brw2020)
```




